Greedy decision trees and tree ensembles often select split variables by
optimizing impurity reduction over many candidate thresholds. This creates a
multiple-comparisons effect: under a complete global null (label permutation invariance
given the covariates), features with more candidate
split points are more likely to appear optimal by chance, producing
selection bias toward high-cardinality features and unstable feature rankings. Conditional
inference trees address this issue by separating variable selection from
split-point optimization via permutation-based hypothesis testing.

We study conditional inference trees and forests that attach permutation-based
hypothesis tests to the variable-selection step at a node (Stage~A; variable
screening), and treat split-point selection (Stage~B; threshold optimization) as
a downstream optimization step.
We focus on the part of the procedure with standard validity guarantees: Stage~A
screening at a fixed (pre-specified) node (especially the root) in fixed-$B$
mode using Monte Carlo permutation p-values with the Phipson--Smyth (+1)
correction. Our inferential target is the nodewise \emph{complete} (global)
permutation null at the node: label exchangeability given the node covariates
and label-independent algorithmic randomness.
Under this exchangeability assumption, the +1 Monte Carlo permutation p-value
is super-uniform (conservative: $\Pr(p \le \alpha) \le \alpha$ under the null),
and with Bonferroni correction we obtain finite-sample bounds on (i) the
probability that Stage~A triggers any split under the complete null and (ii)
the probability of splitting on any particular tested feature. For motivation,
we also give an idealized lemma showing that CART-style greedy selection chooses
features in proportion to their number of candidate splits under an exchangeable
null model (Appendix~\ref{app:cart-bias}).

To support large-scale empirical evaluation, we describe practical
implementation optimizations (e.g., approximate candidate sets and early-stopped
permutation tests) and a two-stage rank-then-evaluate benchmark protocol (rank
features, then evaluate downstream performance on the top-$k$ subset) on
synthetic ground-truth datasets and real tabular datasets. For reproducibility,
we provide a reference implementation. We explicitly distinguish fixed-node
Stage~A p-values (calibrated in fixed-$B$ mode under the complete null) from
adaptive quantities (Stage~B and internal nodes), which are reported as
algorithmic statistics unless additional selective-inference machinery is used
(see \cref{tab:pvalue-scope} for a compact scope summary).
