\section{Introduction}
\label{sec:introduction}

Feature selection in tabular learning serves at least two distinct goals:
improving downstream prediction by removing noise variables, and producing a
ranking that domain users can inspect, audit, and compare across datasets.
Tree-based methods are attractive for both goals because they can capture
nonlinearity and interactions while yielding an embedded notion of variable
importance through the split structure. But that same recursive search makes
statistical interpretation difficult: the chosen variable at a node is the
winner of many data-dependent comparisons, and later nodes are conditioned on
earlier random splits.

This tension is already visible at the root. Classical CART-style trees choose
a variable and threshold by optimizing an impurity objective over a large set
of candidate cutpoints \citep{Breiman1984ClassificationAndRegressionTrees}. For
continuous or high-cardinality features, the number of candidate thresholds can
be much larger than for coarse or categorical features, so the greedy search
implicitly favors variables with more opportunities to attain an extreme score
by chance. This selection mechanism has long been recognized as a source of
bias in tree-based methods and random-forest importance measures
\citep{LohShih1997SplitSelectionMethods,KimLoh2001ClassificationTreesUnbiasedMultiwaySplits,Loh2002RegressionTreesUnbiasedVariableSelectionInteractionDetection,Strobl2007BiasRFVariableImportance}.

Conditional inference trees were introduced precisely to break that coupling
\citep{Hothorn2006UnbiasedRecursivePartitioning}. Instead of searching over all
split thresholds at once, they first screen features for association with the
response via hypothesis testing (Stage~A), and only then optimize a threshold
within the selected feature (Stage~B). That separation is statistically useful
because it creates a well-defined fixed-node screening object on which finite-sample
permutation arguments can be stated. It is also practically useful because the
same Stage~A machinery can be embedded inside trees and forests to induce a
feature ranking.

The difficulty is that this screening object is not the whole algorithm.
Modern tree software is expected to behave like an engineering system: it must
scale to many features, support forests, and keep runtime competitive with
high-throughput baselines such as random forests, extra trees, and boosted-tree
systems \citep{Breiman2001RandomForests,Geurts2006ExtraTrees,ChenGuestrin2016XGBoost,Ke2017LightGBM,Prokhorenkova2018CatBoost}.
Practical implementations therefore add adaptive stopping, candidate-ordering
heuristics, threshold subsampling, and bootstrap-based ensemble randomness.
These choices are often sensible computationally, but they alter the inferential
object. A paper on conditional inference methods has to distinguish clearly
between theorem-backed nodewise screening and algorithmic heuristics used in a
full tree learner.

\paragraph{Two roles: screening vs.\ embedding.}
This manuscript studies Stage~A in both roles. First, we treat fixed-node
Stage~A screening, especially at the root, as a standalone permutation-based
feature screening rule. In that role, the target is a nodewise complete-null
decision event, and the theorem-backed object is the fixed-$B$ Monte Carlo
permutation p-value. Second, we study Stage~A as a component inside a recursive
tree/forest feature ranker. In that role, the tree is an adaptive embedding
procedure, and later-node quantities are best viewed as algorithmic statistics
unless additional selective-inference machinery is introduced. This distinction
is central to the paper and is summarized explicitly in \cref{sec:pvalue-scope}.

\paragraph{Two operating regimes in the implementation.}
The software therefore supports two conceptually different regimes. The
\emph{reference regime} disables early stopping and uses an exhaustive fixed-$B$
Stage~A scan at a fixed node; this is the regime analyzed in
\cref{sec:theory}. The \emph{default engineering regime} enables adaptive
stopping and additional heuristics such as candidate scanning, feature muting,
threshold subsampling, and forest bootstrap resampling. Those behaviors are
important for practical ranking performance and runtime, but they are not all
covered by the fixed-node theorem chain.

\paragraph{What this paper is and is not.}
This is not a general post-selection inference paper for fully adaptive trees,
nor a global false-discovery-control paper for arbitrary feature-selection
workflows. Those are different inferential targets with their own literatures,
including selective inference after model selection
\citep{Berk2013ValidPostSelectionInference,Lee2016ExactPostSelectionInferenceLasso,LeebPotscher2015ConfidenceIntervalsAfterModelSelection},
stability-based variable selection \citep{MeinshausenBuhlmann2010StabilitySelection},
and knockoff-style FDR control \citep{BarberCandes2015Knockoffs}. Our target is
more specific: to make precise which parts of a conditional-inference tree
workflow admit classical finite-sample permutation guarantees, and how those
parts interact with an implementation that also contains pragmatic heuristics.

\paragraph{Contributions.}
This paper makes three contributions.
First, it gives finite-sample, fixed-node guarantees for Stage~A screening in a
fixed-$B$ reference regime under a nodewise \emph{complete} permutation null,
including +1 Monte Carlo p-value super-uniformity and Bonferroni-based
complete-null bounds on Stage~A rejections at the node and on the probability
that the node accepts a split on any particular tested feature. These
guarantees do not depend on the number of candidate thresholds available to a
feature.
Second, it presents a software-aligned account of the engineering optimizations
used to scale conditional inference trees and forests, while drawing a hard
line between theorem-backed fixed-node quantities and adaptive/default outputs
that should be interpreted empirically rather than inferentially.
Third, it evaluates the method as a feature ranker under a pre-specified
rank-then-evaluate benchmark on synthetic and real datasets, so the practical
value of the engineering heuristics is assessed where theory is intentionally
not claimed.

\paragraph{Roadmap.}
\cref{sec:background} sharpens the motivation by reviewing selection bias in
greedy trees, the original conditional-inference framework, and the distinct
inferential targets that arise once such screening is embedded in a full tree
learner. \cref{sec:method} defines the Stage~A/Stage~B procedure and the Monte
Carlo p-values used throughout. \cref{sec:theory} states the fixed-node/root
guarantees, with proofs in the appendices. \cref{sec:experiments} evaluates the
method as a feature-ranking procedure and studies the practical behavior of the
engineering heuristics. \cref{sec:discussion} returns to scope, limitations, and
open methodological directions.
