\section{Background and related work}
\label{sec:background}

\paragraph{Why greedy split selection is statistically delicate.}
At a tree node, a split rule compares many candidate variables and, for each
variable, many candidate thresholds. In CART this is a direct optimization of
an impurity objective over the union of all candidates
\citep{Breiman1984ClassificationAndRegressionTrees}. If feature $j$ has
$|C_j|$ admissible thresholds, then the best score for that feature is an
extreme order statistic over $|C_j|$ opportunities. Even under a complete null
where labels are unrelated to covariates, features with larger candidate sets
tend to achieve more favorable impurity improvements by chance. That is the
core mechanism behind the well-known high-cardinality bias in greedy trees.
The issue predates random forests and motivated alternative split-selection
rules such as QUEST, CRUISE, and GUIDE
\citep{LohShih1997SplitSelectionMethods,Shih2004NoteOnSplitSelectionBiasClassificationTrees,KimLoh2001ClassificationTreesUnbiasedMultiwaySplits,Loh2002RegressionTreesUnbiasedVariableSelectionInteractionDetection};
closely related work also studied unbiased split selection directly for
Gini-based classification trees
\citep{StroblBoulesteixAugustin2007UnbiasedSplitSelectionGini}. The same bias
mechanism reappears in the analysis of random-forest importance measures
\citep{Strobl2007BiasRFVariableImportance}.

\paragraph{Conditional inference as a separation principle.}
Conditional inference trees \citep{Hothorn2006UnbiasedRecursivePartitioning}
replace the single greedy optimization with a two-stage procedure. Stage~A asks
which features are associated with the response at the current node; Stage~B
asks where to split within the chosen feature. The original framework uses
linear test statistics with a permutation-statistics foundation due to
\citet{StrasserWeber1999PermutationStatistics}, and is implemented in widely
used software ecosystems such as the original \texttt{coin}/\texttt{party}
framework and later
\texttt{partykit}
\citep{HothornHornikVandeWielZeileis2006LegoSystemConditionalInference,HothornZeileis2015Partykit}.
The important conceptual point is that the separation changes the object being
analyzed: instead of studying the winner of a large impurity search, one can
study a nodewise screening test. In this paper, that is the inferential anchor.

\paragraph{From the original framework to an implementation-oriented one.}
The present library keeps the conditional-inference separation but broadens the
set of Stage~A statistics and practical execution modes. In addition to the
classical conditional-inference perspective, it supports dependence scores such
as mutual information, distance correlation, and randomized dependence
coefficients \citep{Kraskov2004EstimatingMutualInformation,Szekely2007DistanceCorrelation,LopezPaz2013RDC},
multi-selector aggregation via a max-statistic, tree ensembles, and several
runtime-oriented heuristics. This broader implementation is useful for
benchmarking and deployment, but it also makes it essential to distinguish the
fixed-node object that permutation theory calibrates from the adaptive tree that
practitioners actually run.

\paragraph{Which inferential target is under discussion?}
Several distinct targets are easy to conflate.
One target is \emph{fixed-node screening}: at a pre-specified node, test
whether any of the currently tested features is associated with the response.
That is the target addressed by the theory in this paper.
A second target is \emph{post-selection inference} for the selected feature or
selected threshold after the tree has already adapted to the same data. A third
target is \emph{global feature-selection control} for an end-to-end ranking or
subset-selection workflow.
These latter targets are substantially harder and belong to different
literatures, including selective inference after model selection
\citep{Berk2013ValidPostSelectionInference,Lee2016ExactPostSelectionInferenceLasso,LeebPotscher2015ConfidenceIntervalsAfterModelSelection},
false-discovery control \citep{BenjaminiHochberg1995FDR,BarberCandes2015Knockoffs},
and stability-based selection \citep{MeinshausenBuhlmann2010StabilitySelection}.
We cite them here not to claim those guarantees, but to make clear that our
paper addresses a narrower inferential object.

\paragraph{Multiplicity correction at a node.}
Once Stage~A is viewed as a family of feature-wise tests at a fixed node,
classical multiple-testing tools become relevant. Bonferroni is conservative
but assumption-light and yields immediate complete-null control for the nodewise
decision event. Resampling-based alternatives such as Westfall--Young max-$T$
\citep{WestfallYoung1993ResamplingBasedMultipleTesting} can be less
conservative when dependence across tested features is exploited. In this
manuscript, Bonferroni is the main theorem-backed correction across features,
while multi-selector aggregation within a feature uses the same max-statistic
principle in a narrower setting.

\paragraph{Engineering pressure in modern tree systems.}
The original statistical motivation for conditional inference sits inside a
broader ecosystem of tree learners optimized for predictive performance and
throughput. Random forests and extra trees introduce bootstrap resampling,
feature subsampling, and threshold randomization
\citep{Breiman2001RandomForests,Geurts2006ExtraTrees}; modern boosting systems
use approximate split search and histogram methods at scale
\citep{Friedman2001GBM,ChenGuestrin2016XGBoost,Ke2017LightGBM,Prokhorenkova2018CatBoost}.
These developments matter for this paper because they explain why a practical
conditional-inference implementation accumulates heuristics such as early
stopping, scan ordering, threshold subsampling, and forest resampling. Those
choices are central to runtime and predictive utility, but they should not be
silently conflated with the fixed-$B$ permutation regime.

\paragraph{Feature-selection baselines and evaluation target.}
The broader feature-selection literature spans simple filters, redundancy-aware
filters, wrappers, and embedded methods. Relevant comparators include
correlation- or dependence-based filters, mRMR \citep{Peng2005MRMR}, recursive
feature elimination \citep{Guyon2002RFE}, Boruta \citep{Kursa2010Boruta},
random-forest-based variable-selection heuristics
\citep{GenuerPoggiTuleauMalot2010VariableSelectionUsingRandomForests}, and
permutation-importance variants, including conditional-permutation ideas
\citep{Breiman2001RandomForests,Strobl2008ConditionalVIM,DebeerStrobl2020ConditionalPermutationImportanceRevisited,Pedregosa2011ScikitLearn}.
Because different methods optimize different internal objectives, we evaluate
them primarily as \emph{rankers}: each method outputs an ordering of features,
and that ordering is judged through a downstream rank-then-evaluate protocol
rather than by in-sample impurity alone.

\paragraph{Permutation testing and Monte Carlo p-values.}
Permutation tests provide a nonparametric route to calibration when the labels
are exchangeable under the null relative to the permutation scheme. In the
Monte Carlo setting, the exactness logic behind permutation reference
distributions is classical \citep{Ernst2004PermutationMethodsExactInference},
while empirical p-value estimation from random Monte Carlo draws has its own
calibration issues \citep{NorthCurtisSham2002EmpiricalPValuesMonteCarlo}. When
only random permutations are used, exactness requires care; recent treatments
make precise when identity-plus-random-permutation schemes retain finite-sample
validity \citep{HemerikGoeman2018ExactTestingRandomPermutations}. The
Phipson--Smyth ``+1'' correction then yields finite-sample super-uniform
p-values under exchangeability
\citep{PhipsonSmyth2010PermutationPValues}. There is also a substantial
literature on sequential Monte Carlo testing and bounded resampling risk
\citep{BesagClifford1991SequentialMonteCarloPValues,Gandy2009SequentialImplementationMonteCarloTests}.
Our implementation includes early-stopping heuristics for speed, but this paper
does not invoke that literature to claim optional-stopping validity for the
default engineering outputs.
