\section{Tree-growing method}
\label{sec:method}

At each internal node, conditional inference trees make two decisions. Stage~A
tests candidate features for association with the response and chooses a feature
to split. Stage~B then searches thresholds only for that selected feature; if no
valid split is accepted, the node becomes terminal. In this paper, CIT
denotes the resulting single-tree feature selection method. CIF denotes a
bootstrap ensemble of such trees with candidate feature sampling at each node
and feature rankings formed by aggregating split importances across trees.
CIF-all is the no-candidate-subsampling variant: every nonconstant available
feature can enter Stage~A at each node. This changes the Stage~A feature set,
the Stage~A multiplicity correction, and the per-node computational cost.

\subsection{Stage~A and Stage~B}
\label{sec:stageA-stageB}

At a node $t$, write $(X_t,Y_t)$ for the samples reaching that node. Let
$F_{t,\mathrm{avail}}$ be the feature set carried into the node by the
tree-growing algorithm, including any feature muting inherited from ancestors.
Let $F_{t,\mathrm{nonconst}} \subseteq F_{t,\mathrm{avail}}$ be the features
that are nonconstant on the node samples. After any candidate feature
subsampling, Stage~A considers a set
$F_t \subseteq F_{t,\mathrm{nonconst}}$; its size $m_t=|F_t|$ is the
Bonferroni denominator for that node.
Any label-independent auxiliary randomness used by a selector is denoted by $U$
and is held fixed in the fixed-node guarantee.

When adaptive stopping is enabled during tree growing, Stage~A can evaluate
features sequentially and stop before every feature in $F_t$ has been tested.
Let $E_t \subseteq F_t$ denote the evaluated subset; in the exhaustive fixed-$B$
reference rule, $E_t=F_t$. The fixed-node theorem in \cref{sec:theory} applies
to that reference rule: the features in $F_t$ and the budget $B$ are fixed
independently of the node responses, and every $p_{t,j}$ for $j\in F_t$ uses
that budget. It does not cover response-dependent node selection, inherited
feature muting, or adaptive stopping.

\paragraph{Stage~A (feature selection).}
In the exhaustive fixed-$B$ setting, Stage~A follows the feature-selection and
threshold-search separation of conditional inference trees
\citep{Hothorn2006UnbiasedRecursivePartitioning}. The original conditional
inference tests use the permutation-statistic framework of
\citet{StrasserWeber1999PermutationStatistics}; here, Monte Carlo label
permutations calibrate the configured selector statistic. Stage~A
computes a selector statistic $T^{\mathrm{sel}}_j(X_{t,j},Y_t;U)$ and
permutation $p$-value $p_{t,j}$ for every feature $j \in F_t$, then selects
%
\begin{equation}
  p_t^{\star,\mathrm{ref}} := \min_{j \in F_t} p_{t,j},
  \qquad
  M_t^\star := \{j \in F_t : p_{t,j}=p_t^{\star,\mathrm{ref}}\}.
\end{equation}
%
Let $\tau_t=\alpha_{\mathrm{sel}}/m_t$ when Bonferroni is enabled and
$\tau_t=\alpha_{\mathrm{sel}}$ otherwise. If
$p_t^{\star,\mathrm{ref}} \ge \tau_t$, Stage~A returns no feature and the node
becomes terminal. If $p_t^{\star,\mathrm{ref}} < \tau_t$, the selected feature
$j_t^{\star,\mathrm{ref}}$ is drawn uniformly from $M_t^\star$ using reservoir
sampling, and the node enters Stage~B.

During adaptive tree growing, Stage~A uses the same decision structure but can
order features by a preliminary score and evaluate only the subset $E_t$. The
returned value $q_t^{\mathrm{feat}}$ is used to grow the tree; it is not an
exhaustive fixed-$B$ $p$-value covered by \cref{thm:plusone-superuniform}. Under
the exhaustive reference rule with Bonferroni correction,
$\min(1,m_t p_t^{\star,\mathrm{ref}})$ is a conservative complete-null
$p$-value for the fixed-node Stage~A test. When feature muting is enabled,
features with non-rejecting Stage~A $p$-values can be removed from descendant
candidate feature pools. When muting is disabled, descendants keep
$F_{t,\mathrm{avail}}$ except for deterministic constant-feature pruning.

\paragraph{Stage~B (threshold search).}
Given a selected feature $j_t^\star$, Stage~B starts from midpoints between
ordered unique values of $X_{t,j_t^\star}$. The exact method uses all such
midpoints. The random method samples a bounded subset of midpoints, while the
percentile and histogram methods choose bounded representative thresholds
derived from the midpoint distribution. Candidate thresholds that would leave
either child below the minimum leaf size are filtered out. If no candidate
remains, Stage~B returns no split.

For the remaining finite threshold set $C_{t,j_t^\star}$, Stage~B selects the
threshold used by the tree. The tested statistic at threshold $c$ is the
weighted child impurity
\[
  \frac{|Y_t^L(c)|}{|Y_t|}\mathrm{Imp}(Y_t^L(c))
  + \frac{|Y_t^R(c)|}{|Y_t|}\mathrm{Imp}(Y_t^R(c)),
\]
where $\mathrm{Imp}$ is the chosen node impurity measure. This statistic is used
in a left-tail test, and the exhaustive fixed-$B$ rule chooses the candidate
with the smallest Stage~B $p$-value, with uniform tie breaking. During adaptive
tree growing, threshold scanning may reorder candidates and stop after the first
rejecting threshold, so the returned Stage~B score is a tree-construction score
rather than a global minimum over all retained thresholds. Let
$\tau_t^{\mathrm{split}}=\alpha_{\mathrm{split}}/|C_{t,j_t^\star}|$ when
Bonferroni is enabled for Stage~B and
$\tau_t^{\mathrm{split}}=\alpha_{\mathrm{split}}$ otherwise. If no evaluated
Stage~B $p$-value is below $\tau_t^{\mathrm{split}}$, the node becomes terminal.
After a threshold passes the Stage~B test, the final minimum required-improvement
check uses the corresponding weighted impurity decrease. Stage~B chooses the
split used by the tree; the returned Stage~B score is not a valid post-selection
$p$-value.

\begin{algorithm}[H]
  \caption{Conditional inference tree training skeleton (Stage~A $\rightarrow$ Stage~B)}
  \label{alg:citree}
  \begin{algorithmic}[1]
    \Require Training data $(X,Y)$, hyperparameters $\theta$
    \Ensure Trained tree $\mathcal{T}$
    \State $p \gets$ number of columns of $X$
    \State \Return \Call{GrowNode}{$X,Y,1,\{1,\dots,p\},\theta$}
    \Function{GrowNode}{$X_t,Y_t,d,F_{\mathrm{avail}},\theta$}
      \If{\Call{Stop}{$X_t,Y_t,d,\theta$}}
        \State \Return \Call{Leaf}{$Y_t$}
      \EndIf
      \State $(j_t^\star, q_t^{\mathrm{feat}}, m_t, F'_{\mathrm{avail}}) \gets \Call{StageA}{X_t,Y_t,F_{\mathrm{avail}},\theta}$
      \If{$j_t^\star = \bot$}
        \State \Return \Call{Leaf}{$Y_t$}
      \EndIf
      \State $(c_t^\star, q_t^{\mathrm{split}}) \gets \Call{StageB}{X_t,Y_t,j_t^\star,\theta}$
      \If{$c_t^\star = \bot$}
        \State \Return \Call{Leaf}{$Y_t$}
      \EndIf
      \State $(X_t^L,Y_t^L,X_t^R,Y_t^R) \gets$ \Call{Partition}{$X_t,Y_t,j_t^\star,c_t^\star$}
      \State $\Delta_t \gets \Call{ImpurityDecrease}{Y_t,Y_t^L,Y_t^R,\theta}$
      \If{$\Delta_t$ is below the minimum required impurity decrease}
        \State \Return \Call{Leaf}{$Y_t$}
      \EndIf
      \State Add $\Delta_t$ to the feature-importance accumulator for $j_t^\star$
      \State $u \gets$ internal node with split $(j_t^\star,c_t^\star)$ and stored p-value fields $(q_t^{\mathrm{feat}},q_t^{\mathrm{split}})$
      \State $u.\mathrm{left} \gets \Call{GrowNode}{X_t^L,Y_t^L,d+1,F'_{\mathrm{avail}},\theta}$
      \State $u.\mathrm{right} \gets \Call{GrowNode}{X_t^R,Y_t^R,d+1,F'_{\mathrm{avail}},\theta}$
      \State \Return $u$
    \EndFunction
  \end{algorithmic}
\end{algorithm}

For the benchmarked CIT and CIF configurations, \cref{alg:citree} stores two
values at each split: the Stage~A permutation $p$-value for the selected feature
and the Stage~B permutation $p$-value for the selected threshold.

\subsection{Monte Carlo permutation \texorpdfstring{$p$}{p}-values}
\label{sec:plusone}

Let $T_{\mathrm{obs}}$ be the observed statistic and $T_1,\dots,T_B$ the same statistic
computed on $B$ random label permutations. For a right tail test we use the
Phipson--Smyth $+1$ $p$-value
\citep{NorthCurtisSham2002EmpiricalPValuesMonteCarlo,PhipsonSmyth2010PermutationPValues},
where $\ind\{\cdot\}$ denotes the indicator function:
%
\begin{equation}
  p_{\mathrm{MC}}^{\ge} := \frac{1 + \sum_{b=1}^B \ind\{T_b \ge T_{\mathrm{obs}}\}}{B+1}.
\end{equation}
%
For a left tail test we analogously define
%
\begin{equation}
  p_{\mathrm{MC}}^{\le} := \frac{1 + \sum_{b=1}^B \ind\{T_b \le T_{\mathrm{obs}}\}}{B+1}.
\end{equation}
%
Stage~A uses the right tail convention and Stage~B the left tail convention;
ties count as exceedances in both cases.

\subsection{Forests and runtime hyperparameters}
\label{sec:bootstrap}

CIF combines bootstrap trees and averages their predictions. For feature
selection, it sums each fitted tree's normalized split importances and
renormalizes the forest vector. Each CIF tree uses the same Stage~A and Stage~B
node rules as CIT
\citep{Breiman2001RandomForests,HothornBuehlmannDudoitMolinaroVanDerLaan2006SurvivalEnsembles}.
The fixed-node theorem applies only to the Stage~A reference test described
above.
Across the CIT/CIF benchmark configurations, we hold Bonferroni correction,
adaptive stopping, feature muting, feature scanning, threshold scanning,
histogram-256 thresholding, and CIF bootstrap sampling fixed. The varying
settings are selector choice and honesty.

Runtime depends on the realized work at each node: features tested in Stage~A,
thresholds generated and retained in Stage~B, permutations evaluated before
stopping, and features carried into the node. Adaptive stopping changes the
number of permutations evaluated for each test
\citep{BesagClifford1991SequentialMonteCarloPValues,Gandy2009SequentialImplementationMonteCarloTests}.
Histogram thresholding reduces the Stage~B threshold set by replacing exhaustive
cutpoints with binned thresholds. Feature and threshold scanning reorder
candidates under early stopping, and feature muting can reduce the feature pool
inherited by descendants. Appendix~\ref{app:methods} gives algorithm and
reproducibility details, Appendix~\ref{app:complexity-details} gives runtime and
memory accounting, and \cref{sec:results-practical} reports the measured timing
effects.

\subsection{Feature rankings}
\label{sec:rank-construction}

The experiments evaluate each method through the ranking it produces on the
training fold. CIT and CIF rankings are fitted split-importance rankings, not
rankings of Stage~A $p$-values. For a fitted CIT, feature $j$ receives the sum,
over internal nodes that split on $j$, of the impurity decrease at that node:
parent impurity minus the weighted child impurity. If the total importance is
positive, the vector is normalized to sum to one; if the tree contributes no
positive impurity decrease, the vector remains zero. CIF sums these tree
importance vectors over trees and then normalizes the forest vector in the same
way. Stage~A
determines which features enter threshold optimization, while impurity decrease
scores only the accepted splits by their local reduction in impurity. Larger
importances are ranked earlier, and features never used by a split receive zero
importance.

This split-importance ranking differs from standard random-forest impurity
importance in how a feature enters threshold search. In CART-style trees, the
feature and threshold are chosen together by an impurity search, so a feature
with more candidate thresholds has more chances to produce a large apparent
impurity decrease \citep{Strobl2007BiasRFVariableImportance}. In CIT and CIF,
Stage~A first selects a feature by a permutation test; Stage~B then searches
thresholds only for that feature. This reduces the advantage that
high-cardinality features get from having more possible thresholds. We use these
fitted split-importance scores as rankings and evaluate them through downstream
top-$k$ prediction and synthetic feature recovery.
