\section{Fixed-node Stage~A guarantees}
\label{sec:theory}

This section isolates the exhaustive fixed-$B$ Stage~A feature selection rule
at a fixed node. The theorem proves superuniformity for one $+1$ Monte Carlo
permutation $p$-value, and the corollary applies Bonferroni correction across
the fixed-node Stage~A tests. In the notation of \cref{sec:stageA-stageB}, this is
the exhaustive reference case in which every feature in $F_t$ is tested with the
same effective nodewise fixed budget $B$. All probabilities below are
conditional on $(X_t,U)$, where $U$ denotes label-independent auxiliary
randomness held fixed when conditioning, excluding the Monte Carlo permutation
draws.

\begin{theorem}[$+1$ permutation $p$-values are superuniform]
\label{thm:plusone-superuniform}
Assume A0.1 through A0.5 from \cref{app:assumptions}. Under the nodewise
complete permutation null at the fixed node $t$, conditional on $(X_t,U)$, the
usual computation from the observed statistic and $B$ random label permutations
has the same rank distribution as an exchangeable $(B+1)$-tuple. Define the
right tail $+1$ Monte Carlo $p$-value
%
\begin{equation}
  p_{\mathrm{MC}}^{\ge}
  := \frac{1 + \sum_{b=1}^B \ind\{T_b \ge T_{\mathrm{obs}}\}}{B+1}.
\end{equation}
%
Then for every $\alpha \in [0,1]$,
%
\begin{equation}
  \PP(p_{\mathrm{MC}}^{\ge} \le \alpha \mid X_t,U) \le \alpha.
\end{equation}
\end{theorem}

This is the standard exchangeable rank result for Monte Carlo permutation
tests \citep{NorthCurtisSham2002EmpiricalPValuesMonteCarlo,Ernst2004PermutationMethodsExactInference,HemerikGoeman2018ExactTestingRandomPermutations}; \cref{app:permutation-exchangeability,app:proof-plusone} give the
exchangeability statement and proof. The left tail version follows by
applying the same result to $-T$.

\begin{corollary}[Stage~A complete null control at a fixed node]
\label{cor:stageA-global-null}
Assume A0.1 through A0.5 from \cref{app:assumptions}. Under the nodewise complete null
at $t$, and for a nonempty tested feature set $F_t$ with $m_t = |F_t| \ge 1$,
let Stage~A reject when
%
\begin{equation}
  \min_{j \in F_t} p_{t,j} \le \alpha_{\mathrm{sel}}/m_t.
\end{equation}
%
The strict comparison used by the tree-growing rule defines a subset of this
nonstrict rejection event, so it is at least as conservative.
%
Then
%
\begin{equation}
  \PP\!\left(\exists j \in F_t : p_{t,j} \le \alpha_{\mathrm{sel}}/m_t \,\middle|\, X_t,U\right)
  \le \alpha_{\mathrm{sel}}.
\end{equation}
\end{corollary}

This combines the $+1$ $p$-value superuniformity in
\cref{thm:plusone-superuniform} with the Bonferroni union bound over the $m_t$
Stage~A tests. It controls fixed-node Stage~A feature selection under the
complete null. Partial-null familywise error control, $p$-values for the selected
feature, and guarantees for the full adaptive learner are separate inferential
targets.
