\section{Experimental design}
\label{sec:experiments}

The experiments ask whether rankings produced by CIT and CIF place predictive
features early enough for fixed downstream learners. For each training fold, a
method produces a feature ranking, and fixed downstream learners evaluate the
selected features on the held-out fold. We then summarize overall ranks, how
often CIF performs well across datasets, high-dimensional behavior, recovery on
synthetic data with known informative features, CIF ranking ablations, and
feature sampling in forests.

\subsection{Benchmark setup}
\label{sec:experiments-protocol}

The benchmark uses 5 random seeds, with 5-fold cross-validation within each
seed. Classification fits logistic regression (LR), support vector machine
(SVM), and k-nearest neighbors (KNN) as downstream models and reports balanced
accuracy, defined as mean recall across classes. Regression fits ridge
regression, support vector regression (SVR), and KNN and reports $R^2$. The main
tables use the standard values $k \in \{5,10,25,50,100\}$.
Downstream learner settings are fixed. Logistic regression uses balanced class
weights and a 1,000-iteration limit; SVM uses an RBF kernel with unit
regularization, $\gamma=\texttt{scale}$, and balanced class weights; KNN uses
5 distance-weighted neighbors; ridge uses unit penalty strength; and SVR uses an
RBF kernel with unit regularization, $\epsilon=0.1$, and $\gamma=\texttt{scale}$.

For each seed and fold, feature-standardization parameters are estimated on the
training fold and then applied to the held-out fold. The ranking method uses only
the standardized training fold and produces the ordered feature list described in
\cref{sec:rank-construction}. For each $k$, we take the top-$k$ features from
that ranking, refit standardization on the selected training columns, apply the
same transform to the selected held-out columns, fit the downstream model on the
selected training data, and evaluate it on the held-out fold.

The downstream evaluation also records performance using the full feature set,
$k=p$, for every dataset. For datasets with more than 100 features, the separate
high-$p$ analysis adds fixed values $k \in \{150,200,300,500,750,1{,}000\}$
and fractional values
$\lceil 0.25p\rceil,\lceil 0.5p\rceil,\lceil 0.75p\rceil$, keeping only values
no larger than $p$. The main tables average only the standard $k$ values.

\subsection{Benchmark comparisons}
\label{sec:experiments-analysis-plan}

The benchmark averages scores in three steps. First, for each
$(\text{dataset}, \text{downstream model}, k)$ combination, scores are averaged
over folds and seeds. Second, each dataset score averages those fold-and-seed
means over downstream models and over the standard $k$ values available for the
methods in that comparison. Third, the tables average dataset scores over
datasets.

The real-data classification benchmark contains 23 datasets. Comparisons
involving partykit ctree and cforest use the 22-dataset subset that excludes
\nolinkurl{dexter}; comparisons among methods with \nolinkurl{dexter} results
use all 23 datasets. The complete-case classification analysis, used for
bootstrap intervals and omnibus rank testing, contains 14 datasets with all
17 methods, all three downstream learners, and all five standard values of $k$.
Regression contains 8 datasets; the analogous rank summary uses the
6 complete-case datasets with all methods, downstream learners, and standard
$k$ values available.

By-$k$ results are reported separately for $k=5,10,25,50,100$. The corresponding
classification dataset counts are $22, 21, 15, 15, 14$; the regression counts
are $8, 8, 7, 7, 6$. Pairwise CIF comparisons stratified by $k$ use the dataset
counts reported with the corresponding rows. The 14-dataset classification and
6-dataset regression summaries are complete-case analyses, retaining only
datasets with every method, downstream learner, and standard $k$ value.

\subsection{Datasets}
\label{sec:experiments-datasets}

The real-data benchmark includes 23 classification datasets and 8 regression
datasets. The separate high-$p$ analysis restricts attention to the
15 classification datasets and 6 regression datasets with more than 100
features. The full inventories of the real and synthetic datasets are listed in
\Cref{app:dataset-inventory}, which also summarizes the synthetic designs and
their parameter choices.

Feature recovery uses 8 synthetic classification datasets and 8 synthetic
regression datasets with known informative features and controlled noise
structure. A separate study of feature sampling uses $n=250$ and
varies $p \in \{100, 500, 1{,}000\}$ and
$n_{\mathrm{informative}} \in \{1,2,5,10\}$. That study compares CIT, a CART
decision tree (DT), a random threshold tree (RT), CIF, CIF-all, random forests
(RF), and ExtraTrees (ET), using 1,000 trees for forest methods. CIF-all denotes
the CIF variant that evaluates all nonconstant features at each split.

\subsection{Compared methods}
\label{sec:experiments-methods}

The benchmark includes three types of feature selectors: permutation-test
filters, embedded tree or ensemble methods, and wrapper baselines.
Permutation-test filters compute univariate dependence statistics and rank
features by their permutation $p$-values: multiple correlation (MC) and
randomized dependence coefficient (RDC) for classification, and Pearson
correlation (PC), distance correlation (DC), and RDC for regression. The RDC
filter keeps the empirical-CDF random feature map from
\citet{LopezPaz2013RDC} but replaces the canonical-correlation solve with the
maximum absolute pairwise correlation across projected columns. Distance
correlation follows \citet{Szekely2007DistanceCorrelation}.

Embedded methods rank features using tree or ensemble importance scores. These
methods are CART decision trees (DT), randomized trees (RT), CIT, CIF, random
forests (RF) \citep{Breiman2001RandomForests}, ExtraTrees
\citep{Geurts2006ExtraTrees}, XGBoost \citep{ChenGuestrin2016XGBoost},
LightGBM \citep{Ke2017LightGBM}, CatBoost
\citep{Prokhorenkova2018CatBoost}, and the partykit reference implementations
ctree and cforest
\citep{RPartykitPackage,Hothorn2006UnbiasedRecursivePartitioning,HothornZeileis2015Partykit}.
Wrapper baselines score features through auxiliary random-forest fits: Boruta
\citep{Kursa2010Boruta}, permutation importance (PI)
\citep{Breiman2001RandomForests,Strobl2007BiasRFVariableImportance},
conditional permutation importance (CPI)
\citep{DebeerStrobl2020ConditionalPermutationImportanceRevisited}, and
random-forest recursive feature elimination (RF-RFE)
\citep{Guyon2002RFE,Pedregosa2011ScikitLearn}.

In this benchmark, CPI uses stratified conditional permutation with
correlation-based strata. For each target feature, it trains a 100-tree random
forest on a training split and scores a 20\% validation split. It then
identifies validation-set features with absolute correlation at least 0.5 with
the target feature, forms five-bin percentile strata from those conditioning
features, and permutes the target feature within strata. If no conditioning
feature meets the correlation threshold, the target feature is permuted globally.

Ranking rules use each method's natural score direction. Embedded tree and
ensemble baselines rank features by decreasing fitted importance: DT, RT, RF,
ExtraTrees, XGBoost, LightGBM, and CatBoost. Permutation-test filters rank by
increasing $p$-value, with dependence score as the secondary key. The ctree
reference ranks features by split-use counts. For cforest, we use partykit's
unconditional variable-importance scores with one permutation; CIF uses
split-importance rankings from accepted conditional-inference splits. The
cforest results therefore refer to this benchmarked partykit importance setting.
Boruta and recursive feature elimination rank by their fitted rank variables,
with smaller ranks first. PI and CPI rank by decreasing importance; the CPI row
uses the stratified conditional-permutation procedure described above. Ties are
broken by deterministic, method-specific order for reproducibility.

\subsection{Configuration selection}
\label{sec:experiments-repro}

Several methods, including CIT, CIF, XGBoost, LightGBM, ctree, and cforest, were
evaluated under multiple configurations. Permutation-test filters used one configuration per
statistic. For each task, the paper reports one configuration per method,
selected by mean score over real datasets, downstream models, and the standard
values $k \in \{5,10,25,50,100\}$. Methods with only one configuration are
treated as fixed. The selected configuration is then held fixed across datasets for
that task. Because configuration selection and reporting use the same real-data
benchmark, the reported scores, intervals, and rank summaries are descriptive.
As a sensitivity check, we repeat configuration selection in leave-one-dataset-out
(LODO) form for both tasks. In classification, each of the 14 complete-case
datasets is evaluated using configurations chosen from the other 13 datasets. In
regression, each of the 6 complete-case datasets is evaluated using configurations
chosen from the other 5 datasets. The downstream models, $k$ values, and
averaging within each dataset are unchanged.

For the synthetic analysis, configurations are selected separately from the real-data
benchmark. Each method is assigned the configuration with the highest average recovery
of the true informative features over $k = 5, 10, 25, 50,$ and $100$. The
benchmarked CIT and CIF configurations use adaptive sequential stopping, Bonferroni
correction, feature muting, feature scanning, threshold scanning, and 256-bin
histogram thresholding. The benchmarked CIT and CIF variants differ in selector
and honesty; CIF is run as a bootstrap forest.
Each CIF runtime ablation changes one setting.

To isolate which parts of the CIF ranking matter, we run a ranking ablation on
the real-data benchmark datasets. The ablation holds the selected CIF
configuration fixed and changes one component at a time: replacing fitted split
importance with split counts, disabling feature muting, disabling bootstrap
sampling, or reducing the forest to one tree. Each variant is evaluated with the
same downstream learners, folds, seeds, and standard top-$k$ values as the main
benchmark.
