\section{Discussion and limitations}
\label{sec:discussion}

\paragraph{What our p-values mean.}
At a fixed node $t$, permutation tests are defined conditionally on covariates
treated as fixed by the permutation scheme. For a fixed nodewise hypothesis
(in this manuscript, the nodewise \emph{complete} permutation null in A0.2),
when exchangeability holds under the null given $(X_t, U)$ (where $U$ denotes
algorithmic randomness such as feature subsampling; \cref{app:assumptions}), the
fixed-$B$ +1 Monte Carlo permutation p-value is super-uniform (conservative):
under the null it falls below $\alpha$ with probability at most $\alpha$. This
interpretation targets complete-null calibration at a pre-specified node.
Interpreting Stage~A p-values under featurewise partial nulls (some covariates
associated with $Y_t$) requires a different randomization scheme and is not
pursued here.
See \cref{sec:pvalue-scope} for a compact scope summary.

\paragraph{Exchangeability is a design choice.}
If observations are clustered, longitudinal, or otherwise dependent, naive label
permutations are generally invalid because exchangeability fails. In such
settings, any inferential interpretation requires a restricted permutation
scheme matched to the dependence structure (e.g., permuting within clusters or
within blocks). \citrees does not implement restricted permutations by default,
so calibrated interpretation requires a permutation scheme that matches the
data's exchangeability structure.

\paragraph{Conditional vs.\ marginal inference.}
Permutation tests provide conditional inference based on the covariates treated
as fixed by the permutation scheme. This differs from parametric/asymptotic
tests (e.g., classical ANOVA or correlation tests), which can be more powerful
under their modeling assumptions but require stronger distributional
conditions.

\paragraph{What we do (and do not) claim.}
Our clean inferential statements are scoped to fixed-node
(especially root) Stage~A screening in fixed-$B$ mode. Stage~B tests thresholds
after selecting a feature using the same labels, and internal nodes are adaptive
because node membership depends on earlier label-dependent splits. Without
additional selective-inference machinery (post-selection corrections) or sample
splitting, we report Stage~B and internal-node ``p-values'' as algorithmic
statistics rather than calibrated p-values
\citep{Berk2013ValidPostSelectionInference,Lee2016ExactPostSelectionInferenceLasso,Fithian2014OptimalInferenceAfterModelSelection,Dwork2015ReusableHoldout,LeebPotscher2015ConfidenceIntervalsAfterModelSelection}.

\paragraph{Early stopping.}
\citrees supports early stopping as a computational heuristic for Monte Carlo
permutation testing. In this paper we do not claim optional-stopping validity
for early-stopped outputs; if used, early stopping should be supported
empirically (e.g., via calibration studies).

\paragraph{Other error-control frameworks.}
Our results focus on nodewise Type~I error control for Stage~A screening at a
fixed node under a permutation null. Alternative inferential targets for feature
selection include global false discovery rate (FDR) control
\citep{BenjaminiHochberg1995FDR}, knockoff-based FDR control
\citep{BarberCandes2015Knockoffs}, and stability selection
\citep{MeinshausenBuhlmann2010StabilitySelection}. These methods address
different error metrics and typically rely on different assumptions than the
fixed-node permutation-testing setup studied here; we do not pursue them in this
manuscript.

\paragraph{Sequential boundaries beyond our heuristics.}
More classical sequential-testing frameworks include group-sequential boundary
families \citep{OBrienFleming1979MultipleTesting,Pocock1977GroupSequential} and
alpha-spending functions \citep{LanDeMets1983DiscreteSequentialBoundaries}. We
do not develop or claim any such sequential-validity guarantees for early-stopped
Monte Carlo permutation tests in this paper.

\paragraph{Fixed-$B$ resolution and ties.}
In fixed-$B$ Monte Carlo mode, the +1 p-value takes values on the discrete grid
$\{1/(B+1),2/(B+1),\dots,1\}$. Consequently, if a nodewise threshold is set below
$1/(B+1)$ (e.g., via Bonferroni with $\alpha_{\mathrm{sel}}/m_t < 1/(B+1)$),
then rejection is impossible regardless of signal strength. Discreteness also
implies ties can occur (multiple candidates achieving the same minimal p-value);
when ties occur, the implementation must specify a deterministic or randomized
tie-breaking policy.

\paragraph{Design axes and ablations.}
Beyond the fixed-node theory, \citrees is designed as a research harness for
conditional inference trees: it exposes modular choices for selectors,
splitters, threshold candidate generation, and resampling. Some configurations
overlap qualitatively with widely used training heuristics (bagging, random
threshold proposals, histogram grids); see
\cref{app:methods-connections,tab:citrees-connections}. In this manuscript we
treat these connections as orientation only and reserve empirical comparisons
and ablations for \cref{sec:experiments}.

\paragraph{Practitioner guidance (guarantees vs.\ heuristics).}
To preserve the paper's fixed-node Stage~A p-value guarantees, use fixed-$B$
permutation testing (\texttt{early\_stopping\_selector=None}) with a
label-independent tested family (A0.3): avoid label-adaptive modifications of
the tested set at the node of interest (e.g., disable subtree feature muting if
you want to interpret p-values at a pre-specified internal node). Ensure the
Monte Carlo resolution is compatible with the thresholding rule, i.e.,
$\alpha_{\mathrm{sel}}/m_t \ge 1/(B+1)$ at the node.

For performance-oriented runs, early stopping, feature/threshold scanning, and
muting can reduce computation substantially, but they change the inferential
object. In this manuscript we treat the resulting outputs (including
early-stopped p-value estimates, Stage~B threshold tests, and internal-node
tests) as algorithmic statistics and recommend empirical calibration studies if
they are to be interpreted beyond ranking/optimization.
