\section{Conclusion}
\label{sec:conclusion}

This paper studies how CIT and CIF runtime choices affect feature rankings,
downstream prediction, and synthetic feature recovery. For exhaustive fixed-$B$
Stage~A testing at a fixed node, if the tested features and permutation budget
are fixed independently of the node responses, Bonferroni correction controls
the probability of any Stage~A feature rejection under the nodewise complete
permutation null. The benchmarks then evaluate the full tree and forest
rankings. For CIF, adaptive stopping and exact-versus-histogram threshold search
change runtime the most while keeping Bonferroni correction fixed. CIT timing is
less consistent, so the paper separates it from the CIF runtime and downstream
prediction summaries.

On the real-data benchmarks with the selected settings, CIF ranks fourth of 17
classification methods and third of 18 regression methods. These are
within-benchmark rank and win/loss summaries, not validation of a preregistered
configuration. In direct comparisons, CIF has higher average downstream scores
than ctree and cforest across 22 classification datasets and higher scores than
CIT across 23 classification datasets. In the smaller regression benchmark, CIF
is higher than ctree and cforest on 7 of 8 datasets and higher than CIT on 6 of
8.

The CIF ranking ablations show that forest size matters: using one
tree changes the rankings more than changing feature muting or replacing split
importance with split counts. The high-$p$ and
synthetic analyses show two patterns: CIF recovers true features more reliably
within the top-$k$ list than at rank 1, and candidate sampling inside forests can
leave informative features unavailable at many split decisions. The regression
benchmark is smaller and has wider intervals, so those results are less
conclusive. A next step is to test forest variants that consider more candidate
features at each split while keeping
Stage~A selection separate from Stage~B splitting.
