\section{Setup and assumptions}
\label{app:setup-assumptions}

\subsection{Setup and notation}
\label{app:setup}

Let $(X_i, Y_i)_{i=1}^n$ be training data and consider a fixed node $t$ with
index set $I_t$ (size $n_t$). Write $X_t := (X_i)_{i \in I_t}$ and
$Y_t := (Y_i)_{i \in I_t}$.

At node $t$, the algorithm performs two potential hypothesis testing steps:
Stage~A performs feature selection over features $j \in F_t$ by testing
association with $Y_t$, and Stage~B tests candidate thresholds for the selected
feature $j_t^\star$.

\subsubsection*{Notation summary}
\begin{table}[H]
  \centering
  \small
  \setlength{\tabcolsep}{5pt}
  \begin{tabular}{@{}p{0.22\linewidth}p{0.72\linewidth}@{}}
    \toprule
    Symbol & Meaning \\
    \midrule
    $t$ & Node index (held fixed in ``fixed node'' results). \\
    $n,\ p$ & Global sample size and number of features. \\
    $I_t,\ n_t$ & Sample index set at node $t$ and its size $n_t := |I_t|$. \\
    $X_t,\ Y_t$ & Covariates and labels restricted to node $t$. \\
    $F_t,\ m_t$ & Tested feature set at node $t$ and its size $m_t := |F_t|$. \\
    $E_t$ & Feature subset evaluated during adaptive tree growing; in the exhaustive fixed-$B$ reference rule used by the theorem, $E_t=F_t$. \\
    $F_{\mathrm{avail}}$ & Features available when node $t$ is reached. \\
    $F_{t,\mathrm{nonconst}}$ & Features in $F_{\mathrm{avail}}$ that are nonconstant at node $t$. \\
    $F_{\mathrm{avail}}'$ & Feature set passed to descendants after Stage~A and any feature muting. \\
    $T^{\mathrm{sel}}_j$ & Stage~A selector statistic for feature $j$ at node $t$. \\
    $p_{t,j}$ & +1 Monte Carlo permutation p-value for $T^{\mathrm{sel}}_j$ (fixed-$B$ mode). \\
    $p_{\mathrm{MC}}^{\ge},\ p_{\mathrm{MC}}^{\le}$ & Right/left tail +1 Monte Carlo p-values computed from $B$ random permutations. \\
    $\alpha_{\mathrm{sel}}$ & Stage~A nominal significance level (before Bonferroni adjustment). \\
    $\alpha_{\mathrm{split}}$ & Stage~B nominal significance level (before Bonferroni adjustment). \\
    $M_t^\star$ & Features attaining the minimum fixed-$B$ Stage~A p-value:
    $M_t^\star := \{j \in F_t : p_{t,j}=p_t^{\star,\mathrm{ref}}\}$. \\
    $j_t^{\star,\mathrm{ref}},\ p_t^{\star,\mathrm{ref}}$ & Fixed-$B$
    selected feature and minimum p-value:
    $p_t^{\star,\mathrm{ref}} := \min_{j\in F_t} p_{t,j}$;
    $j_t^{\star,\mathrm{ref}}$ is drawn uniformly from $M_t^\star$ using
    reservoir sampling. \\
    $j_t^\star,\ s_t^\star$ & Feature and Stage~A search score returned during
    adaptive tree growing, possibly after evaluating only part of the feature set. \\
    $C_{t,j}$ & Candidate threshold set for feature $j$ at node $t$ (Stage~B). \\
    $c_t^\star,\ s_t^{\mathrm{split}}$ & Best threshold and its Stage~B split score for the selected feature. \\
    $\pi$ & Bootstrap sampling probabilities over training indices. \\
    $B$ & Permutation resample budget (number of random permutations). \\
    $U$ & Label-independent algorithmic randomness (e.g., feature subsampling, threshold subsampling, tie-breaking randomness). \\
    \bottomrule
  \end{tabular}
  \caption{Compact notation used in the fixed-node statements and proof appendices.}
\end{table}

\subsection{Assumptions and theorem scope}
\label{app:assumptions}

We label assumptions to keep statements and proofs aligned. When a result below
is described as ``fixed node,'' the node $t$ is held fixed rather than chosen
adaptively from the data.
Unless stated otherwise, the results in \cref{sec:theory} use the full
fixed-node checklist below: fixed node, complete-null
exchangeability, label-independent Stage~A family, fixed-$B$ p-values, and
conservative tie handling.

\paragraph{A0.1 (Fixed node).}\label{assump:fixed-node}
The node $t$ is fixed, not selected from the data.

\paragraph{A0.2 (Nodewise complete permutation null / conditional exchangeability).}\label{assump:exchangeability}
Under the (nodewise) complete permutation null at the fixed node $t$, $Y_t$ is
exchangeable conditional on $(X_t, U)$, where $U$ denotes label-independent
algorithmic randomness that can affect which hypotheses are tested (e.g.,
\texttt{max\_features}, threshold subsampling). This assumption targets a
complete null at the node: it generally does \emph{not} hold under a
``featurewise'' partial null in which one feature is null while other features
remain associated with $Y_t$. Monte Carlo permutation draws are additional
randomization independent of the data and are not included in $U$.

\paragraph{A0.3 (Label-independent Stage~A tested family).}\label{assump:label-independence}
For Stage~A results, the tested feature family $F_t$ and resample budget $B$ are
measurable functions of $(X_t, U)$ and do \emph{not} depend on the observed
labels $Y_t$ (or on permutation outcomes). The fixed-node guarantee uses the
exhaustive reference case $E_t=F_t$; early-stopped examined sets are outside
the theorem. A separate fixed-feature Stage~B statement would require its
threshold set to be label-independent.

\paragraph{A0.4 (Fixed-$B$ p-values).}\label{assump:fixed-b}
Permutation p-values are computed with a fixed number of resamples $B$ (no
optional stopping).

\paragraph{A0.5 (Conservative tie handling in the Monte Carlo p-value).}\label{assump:ties}
The +1 Monte Carlo p-value uses the conservative tie convention:
ties are counted as $T_b \ge T_{\mathrm{obs}}$ events for right-tail tests and
as $T_b \le T_{\mathrm{obs}}$ events for left-tail tests. Random tie breaking
among equal best features or thresholds is treated separately as
label-independent algorithmic randomness.

\subsection{Result and assumption checklist}
\label{app:assumption-checklist}

\begin{table}[H]
  \centering
  \small
  \begin{tabular}{p{0.52\linewidth} p{0.42\linewidth}}
    \toprule
    Result & Assumptions used \\
    \midrule
    +1 p-value superuniformity (Theorem~\ref{thm:plusone-superuniform}) & A0.1 through A0.5 \\
    Stage~A complete null control (Corollary~\ref{cor:stageA-global-null}) & A0.1 through A0.5 \\
    \bottomrule
  \end{tabular}
  \caption{Assumptions used by the main fixed-node results in \cref{sec:theory}.}
\end{table}
