\section{Algorithmic and reproducibility notes}
\label{app:methods}

This appendix records implementation details needed to interpret the main
results. Section~\ref{sec:method} gives the algorithmic loop, the setup and
proof appendices give the fixed-node results, and
Appendix~\ref{app:complexity-details} gives the detailed complexity formulas.

\subsection{Scope relative to the theory}

The theory covers a narrow setting: exhaustive fixed-$B$ Stage~A feature
selection at a node specified in advance under the complete permutation null. The
benchmarked learner adds several practical controls that are useful
algorithmically but fall outside that theorem:
\begin{itemize}
  \item early stopping,
  \item feature and threshold candidate ordering inside early stopping,
  \item feature muting,
  \item approximate threshold sets,
  \item bootstrap resampling and forest aggregation,
  \item optional honesty.
\end{itemize}
The fixed-node result applies only to exhaustive Stage~A p-values. The remaining
reported Stage~A and split scores are algorithmic quantities.

\subsection{Node implementation choices}

\paragraph{Stage~A.}
At each node, the citrees implementation removes features that are constant on
the node before any \texttt{max\_features} subsampling. If permutation testing
is enabled, Stage~A computes $+1$ Monte Carlo p-values, can order candidates by
a preliminary score before testing, and may stop early
once a sufficiently strong candidate has been found. When feature muting is
enabled, features with nonsignificant Stage~A scores can be removed from
descendant candidate sets within the same subtree. These choices reduce
computation, but they also make the returned score $s_t^\star$ an algorithmic
Stage~A score rather than a full Stage~A p-value over the complete candidate
set.

\paragraph{Stage~B.}
Given the selected feature, the citrees implementation builds threshold
candidates from midpoints between consecutive unique feature values. Exact search uses all such
midpoints. Random sampling chooses a bounded subset of midpoints; percentile and
histogram rules choose bounded representative thresholds derived from the
midpoint distribution, with a fallback for small samples to the full midpoint
set. When sequential stopping is enabled, threshold scanning may then reorder
the chosen thresholds before testing. Because Stage~B is run after Stage~A has
selected a feature using the node response, its returned quantity
$s_t^{\mathrm{split}}$ is read as an algorithmic split score.

\paragraph{Split statistics.}
For permutation split testing, the implementation uses the unweighted sum
of child impurities as the tested statistic and a left tail convention. The
deterministic split search and the final
\texttt{min\_impurity\_decrease} filter instead use the usual weighted impurity
decrease. This distinction matters only for interpreting the Stage~B score; it
does not change the main benchmark comparisons, which are all based on the final
rankings produced by the fitted learners.

\subsection{Permutation budgets and early stopping}

The citrees implementation uses Phipson--Smyth $+1$ Monte Carlo permutation p-values
\citep{PhipsonSmyth2010PermutationPValues} for both stages: right tail in
Stage~A and left tail in Stage~B. The library exposes integer resampling budgets
and three presets:
\begin{itemize}
  \item \texttt{minimum}: the smallest budget that makes rejection attainable on
  the discrete $+1$ p-value scale,
  \item \texttt{auto}: a larger budget intended to stabilize Monte Carlo error
  near the decision threshold,
  \item \texttt{maximum}: a conservative high-precision setting.
\end{itemize}
When Bonferroni is enabled, the effective threshold depends on the number of
tests at the node, so the realized permutation budget at each node increases with
search width.

\paragraph{Sequential stopping.}
The benchmarked defaults use adaptive sequential stopping. The implementation
tracks exceedances inside batches and stops once a Beta--Binomial posterior is
confident that the latent exceedance rate is either below or above the nodewise
threshold. A simpler futility/significance rule is also available. Both change
the stopping time: once a stage stops at a data-dependent time, the returned
value is no longer a classical fixed-$B$ Monte Carlo p-value. This is
why the main paper separates the fixed-node guarantee from the runtime
ablations.

\subsection{Forests, resampling, and honesty}

CIF averages predictions from bootstrap trees. These forests are empirical
extensions of the Stage~A rule, not settings covered by the fixed-node result
itself. The classification implementation also supports
class-conditional bootstrap variants, and both trees and forests optionally
support honest estimation via sample splitting. These settings are useful
algorithmically, but they change the training distribution or the estimation
target and should therefore be read as empirical design choices rather than as
parts of the theorem.

\subsection{Benchmark protocol}

The appendix tables use the same rank-then-evaluate benchmark protocol as
\cref{sec:experiments-protocol}: a method first produces a ranking on the
training split, and downstream models are then fit on the top-$k$ selected
features and evaluated on held-out data. The real dataset benchmark uses
fivefold cross validation over 5 random seeds, standard values
$k \in \{5,10,25,50,100\}$, and larger high-$p$ values of $k$ only for the
separate high-$p$ saturation check. If a method family has multiple settings,
the analysis uses one chosen setting per task, chosen by mean score on the same
benchmark setup reported in the paper.

\subsection{Benchmark outputs and dataset counts}

The analysis uses completed benchmark outputs without imputation. Some benchmark
outputs are unavailable, and some datasets have fewer than five standard values
of $k$ because values greater than $p$ are not evaluated.
Cell-level summaries omit unavailable cells. Common-support analyses retain only
datasets with every method, learner, and standard value of $k$ present. These
rules explain why ctree and cforest classification comparisons use 22 rather
than 23 datasets and why the common-support analyses use 14 classification and 6
regression datasets.
\Cref{tab:implementation-skips} lists the classification seed skips and the
dexter method exclusion; the regression count comes from the same availability
rule, not from additional rows in this table.

\begin{table}[H]
  \centering
  \scriptsize
  \setlength{\tabcolsep}{4pt}
  \begin{tabular}{@{}lll@{}}
    \toprule
    Method/config & Dataset & Seeds \\
    \midrule
    \texttt{ctree} MonteCarlo & \texttt{gisette} & 3 \\
    \texttt{ctree} MonteCarlo & \texttt{isolet} & 2, 3 \\
    \texttt{cit} with selector \texttt{rdc} & \texttt{gisette} & 0, 1, 3 \\
    \texttt{cit} with selector \texttt{rdc} & \texttt{orlraws10P} & 1 \\
    \texttt{ctree}, \texttt{cforest} & \texttt{dexter} & method exclusion \\
    \bottomrule
  \end{tabular}
  \caption{Documented classification seed-level skips and method exclusions.
  Missing outputs are not imputed; this table documents classification omissions
  only.}
  \label{tab:implementation-skips}
\end{table}

\subsection{Randomness and reproducibility}

The implementation is deterministic given the same base seed. Pure Python
permutation routines use local \texttt{numpy} generators, while compiled
parallel routines use seeded streams per resample because generator RNGs
are not uniformly supported in those contexts. This affects how random streams
are constructed in code, but not the substantive reading of the results: the
benchmark is reproducible from the saved artifacts and the reported rankings are
deterministic given the stored configuration and seed schedule.
