\section{Runtime and memory accounting}
\label{app:complexity-details}

This appendix supports the runtime ablations in \cref{sec:results-practical}.
It gives runtime accounting for the benchmarked tree-growing rule. The
accounting treats training cost as a sum over the nodes reached,
the scans and copies performed, and the candidate tests actually run. It is not
well described by a single product
such as depth times feature count times permutation budget, because candidate
feature pools, node sample sizes, threshold counts, and stopping times all vary across
the tree.

Let $n$ be the training sample count and let $p$ be the number of features. At
node $t$, let $n_t$ be the number of samples reaching the node. Using the
notation of \cref{sec:stageA-stageB}, write $a_t=|F_{t,\mathrm{avail}}|$ and
$d_t^{\mathrm{const}}=|F_{t,\mathrm{avail}}\setminus
F_{t,\mathrm{nonconst}}|$. After constant-feature pruning and any
\texttt{max\_features} subsampling, Stage~A considers $F_t$, with
$m_t=|F_t|$. Let $d_t^{\mathrm{mute}}$ be the number of tested Stage~A features
removed from the descendant candidate feature pool by \texttt{feature\_muting};
if feature muting is disabled, $d_t^{\mathrm{mute}}=0$.

Stage~B is reached only when Stage~A returns a selected feature $j_t^\star$:
either the Stage~A permutation feature test rejects or selector permutation
testing is disabled. If Stage~B is reached, let $\tilde{h}_t$ be the number of
candidate thresholds generated for $j_t^\star$ before any
\texttt{min\_samples\_leaf} filtering. Let
$h_t=|C_{t,j_t^\star}|\le \tilde{h}_t$ be the number of valid thresholds that
remain after this filtering; if no filtering is applied, $h_t=\tilde{h}_t$.

Let $C^{\mathrm{sel}}_j(n_t)$ denote the cost of the Stage~A statistic used for
feature $j$ at node size $n_t$, excluding the label-copy and shuffle work
written explicitly below. In multi-selector mode this is the composite max-$T$
statistic cost for one observed or permuted draw. Let
$C^{\mathrm{split}}(n_t)$ denote the cost of one Stage~B impurity statistic after
the threshold mask and child-label slices have been formed; the mask, slice,
label shuffle, and copy overhead is written explicitly below. For the default
MC/Gini classification path and PC/MSE regression path these statistic
evaluations are linear in $n_t$ for a fixed number of classes. RDC is
$O(n_t\log n_t)$ for fixed random-projection and class counts. Distance
correlation and mutual information should be read through
$C^{\mathrm{sel}}_j(n_t)$: distance-correlation paths may be
$O(n_t\log n_t)$ or $O(n_t^2)$ depending on the backend, and mutual information
is delegated to scikit-learn.

\paragraph{Per-node accounting.}
The tree-growing algorithm first checks the node-level stopping conditions, including
whether all labels at the node are equal, and computes a leaf value if the node
stops immediately. This contributes $O(n_t)$ work at every realized node. If the
node is eligible for splitting, the algorithm scans all currently available
features and removes features that are constant on the node. This costs
$O(a_t n_t)$ for the data scan. Updating the array of available feature indices
costs another $O(a_t d_t^{\mathrm{const}})$ for constant-feature removals in the
dense-matrix algorithm, with worst case $O(a_t^2)$ if
many available features are constant at the same node. This pruning occurs before \texttt{max\_features}, so
\texttt{max\_features} caps Stage~A width but not the initial constant-feature
scan.

Let $\mathcal{J}_t$ be the set of Stage~A features that enter the main
selection loop and whose score or permutation test is actually evaluated before
any feature-level early stop. When permutation testing is enabled,
$\mathcal{J}_t$ is the evaluated subset $E_t$ from \cref{sec:stageA-stageB}.
This set does not include the optional feature-scanning prescan, which is
accounted for separately. Let
$b^{\mathrm{sel}}_{t,j}$ be the
number of permutation draws actually used for feature $j$; when permutation
testing is disabled, take $b^{\mathrm{sel}}_{t,j}=0$. The Stage~A testing work
is
\begin{equation}
  S^{\mathrm{A}}_t
  =
  O\!\left(
    \sum_{j \in \mathcal{J}_t}
      \bigl(b^{\mathrm{sel}}_{t,j}+1\bigr)
      \bigl(C^{\mathrm{sel}}_j(n_t)+n_t\bigr)
  \right),
\end{equation}
where the $+1$ accounts for the observed statistic and the $n_t$ term accounts
for label copies, shuffles, and other per-draw linear overhead inside the
permutation test. Feature scanning, when enabled with selector early stopping,
computes prescan scores and sorts candidates, adding
\begin{equation}
  G^{\mathrm{A}}_t
  =
  O\!\left(
    \sum_{j \in F_t} C^{\mathrm{scan}}_j(n_t)
    + m_t \log m_t
  \right),
\end{equation}
where $C^{\mathrm{scan}}_j$ is the selector-score cost used for ordering. If
\texttt{feature\_muting} removes tested features with non-rejecting Stage~A
$p$-values, updating the descendant array of available feature indices adds
$O(a_t d_t^{\mathrm{mute}})$ current-node update work.

Stage~B candidate construction begins with unique-value and threshold-generation
work on the selected feature. A safe upper bound is $O(n_t\log n_t)$, including
the unique/sort work performed before bounded threshold methods reduce the
returned candidate set. If \texttt{min\_samples\_leaf} is greater than one,
filtering adds $O(n_t\log n_t+\tilde{h}_t\log n_t)$ for the feature sort and
threshold lookups. Bounded non-exact methods reduce $\tilde{h}_t$, but they do
not remove the initial unique-value pass.

Let $\mathcal{K}_t \subseteq C_{t,j_t^\star}$ be the set of retained Stage~B
thresholds that enter the main split-selection loop and whose permutation test
or non-permutation split score is actually evaluated before any threshold-level
early stop. Let
$b^{\mathrm{split}}_{t,k}$ be the corresponding realized permutation count;
when splitter permutation testing is disabled, take
$b^{\mathrm{split}}_{t,k}=0$. The Stage~B testing work is
\begin{equation}
  S^{\mathrm{B}}_t
  =
  O\!\left(
    \sum_{k \in \mathcal{K}_t}
      \bigl(b^{\mathrm{split}}_{t,k}+1\bigr)
      \bigl(C^{\mathrm{split}}(n_t)+n_t\bigr)
  \right).
\end{equation}
In $S^{\mathrm{B}}_t$, the added $n_t$ term accounts for forming or applying
threshold masks, child-label slices, label copies, and shuffles for each
observed or permuted draw.
Threshold scanning, when enabled with splitter early stopping, uses weighted
child impurity as an ordering score and adds
$O(h_t(C^{\mathrm{split}}(n_t)+n_t) + h_t\log h_t)$. This ordering score is
separate from the Stage~B $p$-value statistic.

Finally, if the Stage~B split is accepted and the impurity decrease is large
enough, the dense-matrix tree builder materializes left and right child
matrices with all $p$ columns. This data copy costs $O(n_t p)$ at that split. Nodes that reach
Stage~A but become leaves because Stage~A or Stage~B does not reject still incur
the work already described. Other terminal leaves can also arise after
constant-feature pruning leaves no available feature, after threshold generation
and \texttt{min\_samples\_leaf} filtering leave no valid threshold, or after an
accepted threshold fails the \texttt{min\_impurity\_decrease} gate. Such leaves
contribute the costs incurred up to the last reached code path, but they do not
allocate child matrices.

\paragraph{Fixed-$B$ and adaptive modes.}
In the exhaustive fixed-$B$ reference setting, the realized Stage~A set
$\mathcal{J}_t$ is the full tested feature set, the realized Stage~B set
$\mathcal{K}_t$ is the full retained threshold set, and the realized permutation
counts are the fixed budgets. When permutation testing is enabled, adaptive
stopping can reduce realized work by shrinking the sets
$\mathcal{J}_t,\mathcal{K}_t$ and the realized counts
$b^{\mathrm{sel}}_{t,j}, b^{\mathrm{split}}_{t,k}$. Feature and threshold
scanning are part of the early-stopping tree-growing path; the
fixed-$B$/no-early-stopping reference path bypasses these prescreens. They are
ordering heuristics: they spend prescan work in order to present promising
candidates earlier to the stopping rule. They can reduce realized runtime when
the ordering shortens the stopping path enough to offset that prescan work, but
they do not improve the worst-case bound because the stopping rule can still
examine all candidates and use the full budget.

\paragraph{From nodes to trees and forests.}
Let $T_t$ denote the applicable node-local cost for node $t$: immediate stopping
work, Stage~A work, Stage~B work, and child copy work as far as the recursive
builder reaches. Tree training cost is the sum over all realized tree nodes, not
only over internal split nodes:
\begin{equation}
  T_{\mathrm{tree}} = \sum_{t \in \mathcal{V}(\mathcal{T})} T_t.
\end{equation}
Here $\mathcal{V}(\mathcal{T})$ and $\mathcal{I}(\mathcal{T})$ are the realized
node and internal split-node sets. Leaves stopped by the initial checks
contribute only the node-level stopping and value computation term, while leaves
that fail after Stage~A or Stage~B contribute the corresponding testing terms.
For a binary tree,
$|\mathcal{V}(\mathcal{T})|=2|\mathcal{I}(\mathcal{T})|+1$, and a depth cap
gives the usual exponential crude bound on node count, but $n_t$, $a_t$, $m_t$,
$\tilde{h}_t$, $h_t$, and the realized permutation counts usually change
substantially with depth.

If \texttt{honesty=True}, the node sum is computed on the structure-building
subsample. Training also pays $O(np)$ for the dense split/estimation partition
and a leaf re-estimation pass. If $e$ is the estimation-sample size and
$\ell_i$ is the fitted-tree path length for estimation sample $i$, the current
path representation gives worst-case routing cost
$O(\sum_{i=1}^{e}\ell_i^2)$, followed by a stored-path tree walk of
$O(\sum_{v\in\mathcal{V}(\mathcal{T})} d_v + e)$, where $d_v$ is node depth.

For a forest of $M$ trees, let $s_r$ be the realized sample count passed to tree
$r$ after bootstrap and any classifier sampling method. Let
$C^{\mathrm{sample}}_r$ denote the sample-index construction cost for tree $r$.
This term covers bootstrap and class-conditional sampling, while $O(s_r p)$
accounts for row selection and per-tree dense validation/copy before tree
growth. The forest fit first validates the full input and then fits the trees.
Write $T_{\mathrm{tree},r}$ for the node-sum cost of tree $r$ computed as above.
The forest fit can be summarized as
\begin{equation}
  T_{\mathrm{forest}}
  =
  O(np)
  + \sum_{r=1}^{M}
      \left(C^{\mathrm{sample}}_r + O(s_r p) + T_{\mathrm{tree},r}
      + \mathbf{1}_{\mathrm{honest}} H_{\mathrm{honest},r}\right)
  + O(Mp)
  + \mathbf{1}_{\mathrm{oob}} T_{\mathrm{oob}},
\end{equation}
where $\mathbf{1}_{\mathrm{oob}}$ is one when OOB scoring is requested and zero
otherwise, and $T_{\mathrm{oob}}$ is the optional post-fit OOB scoring cost,
separate from the tree-building sum. Here $H_{\mathrm{honest},r}$ denotes the
honesty partition and leaf re-estimation cost just described, using the tree-$r$
training sample size.
With \texttt{bootstrap=False}, the realized root size is $n$ and
\texttt{max\_samples} is unused; with bootstrap sampling, \texttt{max\_samples}
is a requested sample size or cap, and classifier sampling methods can make the
realized $s_r$ smaller. If honesty is enabled, the tree-structure root size is
the splitting subsample size, and training also pays for the split/estimation
partition plus a leaf re-estimation pass.

\paragraph{How runtime hyperparameters enter the accounting.}
\begin{itemize}
  \item \textbf{\texttt{max\_features}.} This caps $m_t$ and therefore the
  Stage~A testing width. It does not cap the constant-feature preprocessing pass
  over all $a_t$ available features.
  \item \textbf{\texttt{feature\_muting}.} This can shrink descendant values of
  $a_t$ and $m_t$. It can also add $O(a_t d_t^{\mathrm{mute}})$ current-node
  feature-index updates when tested features with non-rejecting Stage~A
  $p$-values are removed.
  \item \textbf{\texttt{threshold\_method} and \texttt{max\_thresholds}.}
  Non-exact threshold methods use \texttt{max\_thresholds} to reduce the
  returned pre-filter threshold set $\tilde{h}_t$ up to method-specific
  constants; \texttt{exact} and the fallback for features with at most four
  unique values use all midpoints.
  \item \textbf{\texttt{early\_stopping} settings, \texttt{feature\_scanning},
  and \texttt{threshold\_scanning}.} With permutation testing enabled, these are
  realized-work optimizations. They change candidate order and may reduce the
  number of candidates and permutations actually evaluated when the stopping
  path benefits from that ordering, but the worst case remains the exhaustive
  candidate feature pool and full permutation budget. In the collected ablations,
  disabling adaptive stopping gives the clearest CIF timing change within the
  early-stopping and scanning runtime hyperparameters. Among the CIF runtime
  comparisons that keep Bonferroni correction fixed, adaptive stopping and
  bounded threshold search are the two largest CIF timing effects. The
  Bonferroni ablation is separate because it changes the statistical rule.
  \item \textbf{Bonferroni correction and resampling presets.} When Bonferroni
  correction is enabled, the nodewise threshold is divided by the number of
  tests at that stage. Writing $r_{\mathrm{tests}}$ for that test count and
  $\alpha_{\mathrm{stage}}$ for the corresponding selector or splitter alpha,
  string-valued resampling presets recompute the permutation budget at the
  stricter threshold $\alpha_{\mathrm{stage}}/r_{\mathrm{tests}}$: the
  \texttt{minimum} budget scales like
  $r_{\mathrm{tests}}/\alpha_{\mathrm{stage}}$, the \texttt{maximum} budget is
  $\lceil r_{\mathrm{tests}}^2/(4\alpha_{\mathrm{stage}}^2)\rceil$ up to
  rounding, and \texttt{auto} uses the same stricter threshold in its
  normal-quantile formula. For an explicit integer budget, the budget is
  multiplied by $r_{\mathrm{tests}}$.
  \item \textbf{\texttt{max\_samples}, \texttt{n\_estimators}, and
  \texttt{n\_jobs}.} These affect forest scaling in different ways:
  \texttt{max\_samples} affects the requested or capped bootstrap sample size,
  \texttt{n\_estimators} sets the number of tree fits, and \texttt{n\_jobs}
  changes tree-fit scheduling and memory pressure.
\end{itemize}

\paragraph{Memory accounting.}
At a node, feature-index arrays and scanning arrays use $O(a_t+m_t)$ space,
and threshold construction uses $O(n_t+\tilde{h}_t+h_t)$ auxiliary space.
Permutation tests add statistic, mask, label, permutation-output, and
selector-specific workspaces; for example, fixed-budget non-adaptive tests may
store an $O(B)$ vector of permuted statistics, while adaptive paths can keep
constant-size counters apart from the copied or shuffled labels and
selector-specific arrays.

The dense-matrix terms dominate in many fits. Validation materializes a dense
$O(np)$ feature matrix. Each accepted split materializes dense left and right
child matrices with all $p$ columns and combined size $O(n_t p)$. During
depth-first recursion, parent frames retain sibling child arrays while one child
is being grown. If $\mathcal{P}$ is the set of accepted split nodes whose child
matrices are live on the active recursion stack, the dense-array part of peak
single-tree training memory is bounded by
\[
  O\!\left(np + p\sum_{u\in\mathcal{P}} n_u\right),
\]
plus node-local auxiliary workspaces. This is a live-path bound: it is $O(np)$
for balanced growth but can be larger for highly unbalanced recursion.

Honesty adds fit-time memory for the dense split and estimation partitions. It
also stores estimation-sample groups by leaf path, giving
$O(e+\sum_{q\in\mathcal{Q}} |q|)$ path-index workspace for the set
$\mathcal{Q}$ of retained leaf-path keys, bounded by $O(eD)$ when all estimation
samples have path length at most $D$. This is in addition to the tree-building
memory above.

Forest peak training memory grows with concurrent tree fits. Let
$q=\min(M,\texttt{n\_jobs}_{\mathrm{eff}})$ be the number of trees fit
concurrently by the joblib backend. Peak memory includes the validated forest
input, returned estimators, and up to $q$ per-tree workspaces: bootstrap
indices, dense row-selected matrices, validation copies, and tree-local
peak-memory terms.

For the same $M$-tree forest, persistent fitted storage includes the stored node
records over trees, the estimator list, $O(Mp)$ per-tree fitted arrays such as
feature names and feature importances, and the forest-level $O(p)$
feature-importance and metadata arrays. When \texttt{oob\_score=True}, fitted
storage also includes $O(nK)$ classifier OOB decision values for $K$ classes or
$O(n)$ regressor OOB predictions, plus the scalar OOB score. OOB scoring also
uses temporary fit-time workspace for the OOB count vector, recomputed bootstrap
and unsampled indices, prediction slices $X_{\mathrm{oob}}$, and per-tree
predictions or class-probability arrays.
