\section{Implementation notes (aligned to \citrees)}
\label{app:implementation}

This appendix collects implementation details relevant to reproducibility and
to the scope of the theoretical statements in \cref{sec:theory}.

It complements Appendix~\ref{app:methods}: Appendix~\ref{app:methods} gives
the algorithmic details, while this appendix turns to reproducibility
details such as RNG behavior, compilation modes, and implementation choices
that affect how the results should be read.

\subsection{Benchmark tables and selected-config audit}

The paper-facing benchmark summaries in Sections~\ref{sec:experiments} and
\ref{sec:results} are regenerated from the saved experiment artifacts by
\texttt{paper/scripts/analysis/build\_paper\_benchmark\_tables.py}. For the
real-data benchmark, the canonical outputs include:
\begin{itemize}
  \item \texttt{paper\_benchmark\_best\_configs.csv} for the selected method IDs,
  \item \texttt{paper\_benchmark\_selected\_config\_details.csv} for resolved
    parameter settings, candidate grid sizes, and runner-up gaps within each
    method family,
  \item \texttt{paper\_benchmark\_complete\_case\_membership.csv} for the
    changing standard-budget support, and
  \item \texttt{paper\_benchmark\_fixed\_panel\_aggregate.csv} for the stricter
    common-dataset sensitivity summary, and
  \item \texttt{paper\_benchmark\_fixed\_panel\_membership.csv} for the
    underlying fixed-panel dataset inventory.
\end{itemize}
These tables are the reproducibility surface for the benchmark numbers quoted
in the main text.

\subsection{Documented benchmark skips}
\label{app:implementation-skips}

The packaged benchmark includes a small number of documented skips that persist
in the saved artifact tree and therefore affect complete-case support. The main
text refers to them collectively; the concrete inventory is:

\begin{table}[t]
  \centering
  \scriptsize
  \setlength{\tabcolsep}{4pt}
  \begin{tabular}{@{}lll@{}}
    \toprule
    Method/config & Dataset & Seeds \\
    \midrule
    \texttt{r\_ctree} MonteCarlo & \texttt{gisette} & 3 \\
    \texttt{r\_ctree} MonteCarlo & \texttt{isolet} & 2, 3 \\
    \texttt{cit} with selector \texttt{rdc} & \texttt{gisette} & 0, 1, 3 \\
    \texttt{cit} with selector \texttt{rdc} & \texttt{orlraws10P} & 1 \\
    \texttt{r\_ctree}, \texttt{r\_cforest} & \texttt{dexter} & method-level exclusion \\
    \bottomrule
  \end{tabular}
  \caption{Documented persistent skips in the packaged benchmark. These cells
  are retained as missing rather than imputed, so they affect complete-case
  support and fixed-panel membership.}
  \label{tab:implementation-skips}
\end{table}

\subsection{Registry pattern}

\citrees uses a registry mechanism that maps short identifiers for association
statistics and split criteria to their corresponding implementations and
permutation-test adapters. This keeps the codebase extensible while preserving a
small, stable external interface.

\subsection{Testing modes (compiled/uncompiled)}

Performance-critical routines are compiled for speed. The test suite supports
both compiled execution (to validate optimized code paths) and an uncompiled
execution mode (to enable line coverage measurement in CI).

\subsection{Randomness and reproducibility}

\citrees supports reproducible randomness via a user-provided base seed, but the
implementation uses two RNG patterns depending on the execution path.
Non-JIT permutation-test routines initialize a local
\texttt{np.random.default\_rng(random\_state)} stream, while compiled parallel
Numba routines use per-resample seeding inside parallel loops,
%
\begin{equation}
  s_i := s_0 + i,
\end{equation}
%
because generator-based RNG streams are not uniformly supported in those
contexts. The result is deterministic across runs given the same base seed.

At a fixed node, the same base seed is reused across candidate-wise tests
unless the caller changes it. Thus different per-feature or per-threshold tests
are reproducible but not generated from independent candidate-indexed RNG
streams. This is an implementation choice, not a theorem assumption.

\subsection{Bootstrap methods in forests}

Forests can optionally resample training rows per tree. In bootstrap mode, the
implementation uses with-replacement resampling, either via classical uniform
bootstrap or, for classification forests, via class-conditional variants such
as stratified, undersampling, and oversampling bootstraps. These choices affect
the per-tree resampling distribution, and they can invalidate the root-level
exchangeability assumption A0.2 used in \cref{sec:theory} because duplicated
indices induce non-permutation-invariant equality constraints in the bootstrap
sample.
Accordingly, the fixed-node/root validity statements apply only to trees or
nodes for which A0.2 still holds after resampling (for example, full-data roots
or subsampling without replacement).
